### 2021-10-29 manuscript/hiv-gwas.tex

Our premise is that spVL is the sum of three parts: direct host genetic effects, viral genetic effects, and environmental effects. Uniquely, the viral part is heritable from one transmission partner to another \parencite{Mitov2018, Bertels2018}. To characterize these parts, we use a phylogenetic mixed model (PMM, \cite{Housworth2004}). PMMs assume continuous traits like spVL are the sum of independent heritable and non-heritable parts. The heritable viral part is modeled by a random process occurring in continuous time along the branches of the viral phylogeny. The non-heritable host and environmental parts are modeled as gaussian noise added to sampled individuals at the tips of the phylogeny (Figure \ref{fig:spVLModel}).

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{figures_archived/200207_qq_plots.png}
	\caption{Quartile-quartile plots from association tests.}
	\label{fig:qq-plots}
\end{figure}

>>>>

Our premise is that spVL is the sum of several parts: direct host genetic effects, pathogen genetic effects, interaction effects between host and pathogen genetic factors, and environmental effects. Of these, only pathogen genetic effects should be heritable from one transmission partner to another \parencite{Mitov2018, Bertels2018}. To characterize these effects, we use a phylogenetic mixed model (PMM, \cite{Housworth2004}). PMMs assume continuous traits like spVL are the sum of independent heritable and non-heritable parts. In our case, the heritable part comes from pathogen genetic effects and the non-heritable part comes from the other effects. Then, viral genetic effects are modeled by a random process occurring in continuous time along the branches of the pathogen phylogeny. The other, non-heritable effects are modeled as Gaussian noise added to sampled individuals at the tips of the phylogeny (Figure \ref{fig:spVLModel}).

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{figures/qq_plots.png}
	\caption{Quartile-quartile plots from association tests. The dashed green line shows the y = x line.}
	\label{fig:qq-plots}
\end{figure}
